---
title: "Dashboard for HW 4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(flexdashboard)
library(tidyverse)
library(p8105.datasets)
library(plotly)
library(patchwork)
data(instacart)

insta_small = instacart %>% 
  sample_n(500000) %>%
  mutate(order_dow = recode(order_dow, 
       "1"="Sunday",
       "2"="Monday",
       "3"="Tuesday",
       "4"="Wednesday",
       "5"="Thursday",
       "6"="Friday",
       "0"="Saturday"), 
       order_dow = as.factor(order_dow)) %>%
  mutate(order_dow = ordered(order_dow, levels = c("Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday")))


```


Column
-----------------------------------------------------------------------

### Days since last order and order size

```{r, echo=FALSE}
matrix_for_hm = insta_small %>%
  filter(days_since_prior_order<30) %>%
  group_by(days_since_prior_order) %>%
  count(order_id) %>%
  select(-order_id) %>%
  rename(order_size = n) %>%
  group_by(days_since_prior_order, order_size) %>%
  count(order_size) %>%
  pivot_wider(names_from = days_since_prior_order , values_from= n)%>%
  as.matrix()
  plot_ly(z = matrix_for_hm, type = "heatmap") %>%
  layout(xaxis = list(title = "Days since last order"), yaxis = list(title = "Number of Items in order"))
```

### Explanation for heatmap

Among regular instacart users (days since prior order<30 days), is number of days since last order related to size of the order?

One could make a case that within 10 days, those that order after more days tend to order more items, but this small trend may just be driven by many people ordering around 7 days, resulting in a higher frequency of orders containing 5 items in the 7th day than on the 2nd day since ordering simply because there are more people ordering 7 days after their last order. A box plot of the first 10 days on the x axis would be my next investigative step here. I wanted to make a heatmap first because I thought of the variables as continuous. 


Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Order Size
```{r, echo=FALSE}
 insta_small %>%
  group_by(order_dow) %>%
  count(order_id) %>%
  plot_ly(x = ~order_dow, y = ~n, type = "box") %>%
  layout(xaxis = list(title = "Day of the Week"), yaxis = list(title = "Number of items in order"))
```

### Order size text
Does the size of the order vary by day of the week? 

No, the size of the order does not vary much by day of the week. A small case could be made that orders on Fridays tend to be larger.


### Orders Per Day

```{r, message = FALSE, , echo=FALSE}
insta_small %>%
  group_by(order_dow) %>%
  summarize( num_orders= n_distinct(order_id)) %>%
  plot_ly(x = ~order_dow, y = ~num_orders, type = "bar") %>% 
  layout(xaxis = list(title = "Day of the Week"), yaxis = list(title = "Number of Orders"))

```

### Orders per day text

Which day of the week is most popular for ordering from instacart?
(Days of the week converted from number to name with advice on which day is which from this github page:https://gist.github.com/jeremystan/c3b39d947d9b88b3ccff3147dbcf6c6b)

We see that the most orders come in on Saturdays. 

